---
title: "Hypothesis Testing Analysis"
author: "Celumusa Fakudze"
date: "2024-05-18"
output: html_document
---

## Importing dataset into R-studio environment


```{r}
library(readr)
insurance <- read_csv("C:/Users/Fakudze/Desktop/EswatiniUser/datasets/insurance.csv")
View(insurance)
```


## Loading packages 


```{r}
library(tidyverse)
library(dplyr)
library(summarytools)
library(Hmisc)
library(mice)
```


## Checking for any missing values in the dataset


```{r}
sum(is.na(insurance))
```


## Studying data types


```{r}
str(insurance)
```


## Clean smoker variable

The smoker variable responses are recoded as numerics to ease analysis and then converted to a factor.


```{r}
insurance$smoker[insurance$smoker == "no"] <- "1"
insurance$smoker[insurance$smoker == "yes"] <- "2"
insurance$smoker <- as.factor(insurance$smoker)
```


## Create a smoker dummy variable


```{r}
insurance$smoker_dummy <- as.integer(insurance$smoker == "2")
```


## Calculate the smoking rate


```{r}
sum(insurance$smoker_dummy)/nrow(insurance)*100
```


## Hypothesis testing for media claims


### **Claim 1**: Insurers who smoke are charged more than insurers who does not smoke.


```{r}
smokers <- subset(insurance, smoker_dummy == 1)
non_smokers <- subset(insurance, smoker_dummy == 0)
mean(smokers$charges, na.rm = TRUE)
mean(non_smokers$charges, na.rm = TRUE)
```


### Is this statistically significant? A t-test is employed


```{r}
t.test(smokers$charges, non_smokers$charges, alternative = "two.sided", conf.level = 0.95)
```


### **Claim 2**: Sex has no effect to smoking


```{r}
men = subset(insurance, sex == "male")
women = subset(insurance, sex == "female")
mean(men$smoker_dummy, na.rm = TRUE)
mean(women$smoker_dummy, na.rm = TRUE)

```


### Is this staistically significant? Lets put it into a t.test


```{r}
t.test(men$smoker_dummy, women$smoker_dummy, alternative = "two.sided", conf.level = 0.95)
```

